\textbf{Computational Fuzzy Extractors}

### 2013-06-14 CompFuzzExt.tex
Changing \vW to \vQ for consistency

\newcommand{\ve}{\vect{e}}
\newcommand{\vm}{\vect{m}}
\newcommand{\vy}{\vect{y}}
\newcommand{\vE}{\vect{E}}
\newcommand{\vS}{\vect{S}}
\newcommand{\vA}{\vect{A}}
\newcommand{\vc}{\vect{c}}
\newcommand{\vW}{\vect{W}}
\newcommand{\vR}{\vect{R}}
\newcommand{\vU}{\vect{U}}
\newcommand{\vT}{\vect{T}}
\newcommand{\vX}{\vect{X}}
\newcommand{\vB}{\vect{B}}
\newcommand{\vz}{\vect{z}}
\newcommand{\vd}{\vect{d}}
\newcommand{\vs}{\vect{s}}
\newcommand{\vx}{\vect{x}}
\newcommand{\va}{\vect{a}}
\newcommand{\vb}{\vect{b}}
\newcommand{\vgamma}{\mathbf{\Gamma}}
\newcommand{\vt}{\vect{t}}
\newcommand{\vu}{\vect{u}}
\newcommand{\vF}{\vect{F}}
\newcommand{\recout}{x}
\newcommand{\ignore}[1]{}
\newcommand{\M}{\mathcal{M}}

@@

To extract pseudorandom bits, we use a result of Akavia, Goldwasser, and Vaikuntanathan~\cite{akavia2009} to show that $X$ has many hardcore bits.  The result says that if $\distLWE_{(n-k, m, q, \chi)}$ is secure then any $k$ variables of $X$ in a $\distLWE_{(n, m, q, \chi)}$ instance are hardcore.  We state their result for a general error distribution~(noting that their proof does not depend on the error distribution):
\begin{lemma}\protect{\cite[Lemma 2]{akavia2009}}
\label{lem:many hardcore bits}
 If $\distLWE_{(n-k, m, q, \chi)}$ is $(\epsilon, s_{sec})$ secure then \lnote{notation doesn't exactly match def 4.2}
\[\delta^ {\mathcal{D}_{s_{sec'}}} ((X_{1,..., k}, \vA, \vA X+E) , (U, \vA, \vA X+E)) \le \epsilon\,,\]
where $U$ denotes the uniform distribution over $\Fq^m$,  $s_{sec}' \approx s_{sec} - n^3$, and the remaining variables are as in \defref{def:dist-LWE}.
\end{lemma}

@@

\ignore{
We now study \consref{cons:informal construction} with a symbol fixing source.

\begin{assumption}[LWE w/ entropic errors]
Let $n$ be a security parameter and define $m = \poly(n)$ and $q\geq 2$ be integers.  Let $W = W_1||...||W_{m+\alpha}$ be $(m+
\alpha, m, 2^b)$ symbol-fixing source over $\{0,1\}^{(m+\alpha)\ell}$. For a vector 
Let $A\in\Fq^{(m+\alpha)\times (n+\alpha)}$ and let $x\in\Fq^{n+\alpha}$.  No PPT algorithm given $Ax+W$ can recover $x$ with probability greater than nonnegligible.
\end{assumption}
}
We now define the decisional $\LWE$ problem with symbol fixing errors: 
\begin{problem}[$\distLWE$ w/ symbol fixing errors]
Let $n$ be a security parameter.  Let $q=q(n)=\poly(n)$ be a prime and  $m = m(n)=\poly(n)$ be an integer.  Let $W = W_1||...||W_{m+\alpha}$ be an $(m+
\alpha, m, |Z|)$ symbol fixing source over  $Z^{m+\alpha}$ for the alphabet $Z=[-2^{b-1},2^{b-1}]$. For a vector $X \leftarrow \Fq^n$, let $A_{X, W}$ be the distribution obtained by choosing a uniform $\vA\in\Fq^{(m+\alpha)\times (n+\alpha)}$ and outputting $\vA X + W$. The decisional learning with errors problem $\distLWE_{n+\alpha, m+\alpha, q, W}$ withAt symbol fixing errors is to distinguish samples chosen according $A_{X, W}$ \lnote{why is $X$ a subscript in this notation by A is uniform? why not both?}   \lnote{notation doesn't exactly match def 4.2}
for a uniformly random $X \in \Fq^{n+\alpha}$ from samples chosen according to the uniform distribution over $\Fq^{n+\alpha} \times \Fq$.
\end{problem}

The following theorem states the main technical result of this section, which is of potential interest outside our specific setting. The result is that $\distLWE$ with symbol-fixing sources is implied by standard $\distLWE$,  with security degrading exponentially as symbols are fixed (of course, once $n$ symbols are fixed, no security remains, because there are $n$ linear equations with known error and $n$ unknowns).
%\vspace{.1in}
\begin{theorem}
\label{thm:blockLWE}
 \lnote{notation doesn't exactly match def 4.2}
Let $n$ be a security parameter and define $q, m , \alpha = \poly(n)$.  Let $\beta  = \omega(1)$, and $W$ be a $m$-symbol fixing source over $\{0,1\}^{(m+\alpha)b}$~(that is, $\Hoo(W) = mb$).  The $\distLWE_{n, m,q, W}$ assumption implies the $\distLWE_{n+\alpha+\beta, m+\alpha, q, W}$ w/ symbol fixing sources problem.
\end{theorem}

@@

\begin{figure}[p]
\begin{framed}
\begin{enumerate}
\item Input $\vA', \vb'$, where $\vA' \overset{\$} \leftarrow \Fq^{m\times n}$ and $\vb'$ is either uniform over $\Fq^m$ or $\vb' = \vA'\vx' +\ve'$.
\item Randomly generate $\vect{R} \overset{\$}\leftarrow \Fq^{\alpha \times n}$. Initialize $\vW \in \Fq^{n\times (\alpha+\beta)}$ to be the zero matrix.
\item Let $\vb^* = (\vb', b^*_{m+1}, \ldots,b^*_{m+\alpha})$, where $(b^*_{m+1}, \ldots, b^*_{m+\alpha} )\overset{\$} \leftarrow \Fq^\alpha$.\label{step:b generation}
\item Randomly generate $\vect{S} \overset{\$}\leftarrow \Fq^{\alpha \times (\alpha+\beta)}$.
		\subitem If $\rank(\vect{S})<\alpha$, stop and output a random bit.
\item Find a set of $\alpha$ linearly independent columns in $\vS$.  Let $T$ be the set of these indices.
\item For all $x_i$ where $n\leq i$ and $i\not \in T$:\label{step:fill in matrix}
\subitem Generate $x_i\overset{\$}\leftarrow \Fq$.  
\subitem For $j=1,..., m$:
\subsubitem Generate $\vW_{j, i}\overset{\$}\leftarrow \Fq$
\subsubitem Set $b_j^* = b_j^* + \vW_{j, i} x_i$.
\item Initialize $\vA^*  = \left(\begin{array}{c | c}\vA' & \vW\\\hline \vR & \vS\end{array}\right)$.
\item \label{step:randomization}
For {$i=1,..., m$}:
\subitem Randomly generate $\gamma_i \leftarrow \Fq^{1 \times \alpha}$.
\subitem Set $\va_{i} \leftarrow \va^*_{i}+\gamma_i (\vR||\vS)$
\subitem Set $b_i \leftarrow b^*_i + \gamma_i (b^*_{m+1},..., b^*_{m+k})$
\item Run $D$ on input $\vA, \vb$.
\item Output $D$'s output.
\end{enumerate}
\end{framed}
\caption{A PPT $D'$ that distinguishes LWE using distinguisher for LWE w/ block fixing source}
\label{fig:perfectLWEreduction}
\end{figure}
The distinguisher $D'$ has an advantage when $S$ is of rank $\alpha$.  This occurs with overwhelming probability~(techniques from Cooper~\cite{cooper2000rank}):
\begin{claim}
\label{cl:full rank matrix}
Let $\vS \overset{\$}\leftarrow \Fq^{\alpha \times (\alpha+\beta)}$ be randomly generated.  Then $\Pr[\rank(\vS)=\alpha]\geq  1- \ngl(n)$.
\end{claim}
\begin{proof}
The probability that all $\alpha$ rows are linearly independent is the successive probability that each row is linearly independent of the previous rows, for the $i$th row this probability is $q^{i-1}/q^{\alpha+\beta}$.  By union bound and $\beta = \omega(1)$, the probability that $\vS$ is not full rank is:
\begin{align*}
\Pr[\rank(\vS) < \alpha] &= \sum_{i=1}^{\alpha} \frac{q^{i-1}}{q^{\alpha+\beta}}
\leq \alpha \frac{q^{\alpha}}{q^{\alpha+\beta} }=\ngl(n)\\
\end{align*}
Thus, $\Pr[\rank(\vS) = \alpha] \geq 1-\ngl(n)$.
\end{proof}
\begin{corollary}
$
|\Pr[D(\vA, \vA\vx+\ve) = 1  | \rank(\vS) = \alpha] - \Pr[D(\vA, \vu) =1 | \rank(\vS) = \alpha]|> \epsilon - \ngl(n).
$
\end{corollary}

@@

\begin{claim}
\label{cl:randomMatrixDist}
The matrix $\vA$ generated after Step \ref{step:randomization} is a random matrix with $\rank(\vS) = \alpha$.
\end{claim}
\begin{proof}
It should be clear that $\vR, \vS$ are randomly generated~(conditioned on $\rank(\vS) =\alpha$).  The quadrant represented by $\vA'$ remains random after step~\ref{step:randomization} as we are adding uncorrelated random values.  
The matrix $\vW_{T^c}$~(recall this is the restriction of $\vW$ to the columns not in $T$) is random as it is initialized with random values and only uncorrelated random values are added.

Thus, we restrict attention to $\vW_T$~(the columns of $W$ included in $T$).  Consider a particular row $i$.  Then $\vW_{i, T} = \gamma_i \vS_{T}$.  Since $\vS_T$ is full rank and $\gamma_i$ is uniformly generated, $\vW_{i, T}$ is uniformly generated.

@@

Let $\vb^*$ be the vector generated after step~\ref{step:fill in matrix}.  The only correlation between $\vb^*$ and $\vA^*$ is in the matrix $\vW$, i.e. for $1\leq i\leq n$, $b_i^* = b_i + \sum_{n<j \wedge j\not\in T} \vW_{i, j} x_j$, thus conditioning on $\vW$, $b_i^*$ is uniform assuming $b_i$ is uniform.  Thus $\vb^* | \vA^*$ is uniformly distributed. We show that $\vb |( \vA, \vA^*)$ is uniformly distributed, therefore, showing that $\vb | \vA$ is uniformly distributed after step~\ref{step:randomization}.  Let $\vgamma$ represent the matrix formed by $\gamma_{i}$.

>>>>

\newcommand{\ve}{\vect{e}}
\newcommand{\vm}{\vect{m}}
\newcommand{\vy}{\vect{y}}
\newcommand{\vE}{\vect{E}}
\newcommand{\vS}{\vect{S}}
\newcommand{\vA}{\vect{A}}
\newcommand{\vc}{\vect{c}}
\newcommand{\vW}{\vect{W}}
\newcommand{\vQ}{\vect{Q}}
\newcommand{\vR}{\vect{R}}
\newcommand{\vU}{\vect{U}}
\newcommand{\vT}{\vect{T}}
\newcommand{\vX}{\vect{X}}
\newcommand{\vB}{\vect{B}}
\newcommand{\vz}{\vect{z}}
\newcommand{\vd}{\vect{d}}
\newcommand{\vs}{\vect{s}}
\newcommand{\vx}{\vect{x}}
\newcommand{\va}{\vect{a}}
\newcommand{\vb}{\vect{b}}
\newcommand{\vgamma}{\mathbf{\Gamma}}
\newcommand{\vt}{\vect{t}}
\newcommand{\vu}{\vect{u}}
\newcommand{\vF}{\vect{F}}
\newcommand{\recout}{x}
\newcommand{\ignore}[1]{}
\newcommand{\M}{\mathcal{M}}

@@

To extract pseudorandom bits, we use a result of Akavia, Goldwasser, and Vaikuntanathan~\cite{akavia2009} to show that $X$ has many hardcore bits.  The result says that if $\distLWE_{(n-k, m, q, \chi)}$ is secure then any $k$ variables of $X$ in a $\distLWE_{(n, m, q, \chi)}$ instance are hardcore.  We state their result for a general error distribution~(noting that their proof does not depend on the error distribution):
\begin{lemma}\protect{\cite[Lemma 2]{akavia2009}}
\label{lem:many hardcore bits}
 If $\distLWE_{(n-k, m, q, \chi)}$ is $(\epsilon, s_{sec})$ secure then
\[\delta^ {\mathcal{D}_{s_{sec'}}} ((X_{1,\dots, k}, \vA, \vA X+\chi) , (U, \vA, \vA X+\chi)) \le \epsilon\,,\]
where $U$ denotes the uniform distribution over $\Fq^k$,  $\vA$ denotes the uniform distribution over $\Fq^{m\times n}$, $X$ denotes the uniform distribution over $\Fq^n$, $X_{1,\dots, k}$ denote the first $k$ coordinates of $x$, and $s_{sec}' \approx s_{sec} - n^3$.
\end{lemma}

@@

The following theorem states the main technical result of this section, which is of potential interest outside our specific setting. The result is that $\distLWE$ with symbol-fixing sources is implied by standard $\distLWE$ (but for $n$ and $m$ reduced by the amount of fixed symbols).
%\vspace{.1in}
\begin{theorem}
\label{thm:blockLWE}
Let $n$ be a security parameter, $\alpha$ be polynomial in $n$, and $\beta=\beta(n)=\omega(1)$.
Let $U$ denote the uniform distribution over $Z^m$ for an alphabet $Z\subset \Fq$, and let $W$ denote an $(m+\alpha, m, |Z|)$ symbol fixing source over $Z^{m+\alpha}$.
If $\distLWE_{n, m,q, U}$ is secure, then $\distLWE_{n+\alpha+\beta, m+\alpha, q, W}$ is also secure.
\end{theorem}

\thref{thm:blockLWE} holds for an arbitrary error distribution~(not just uniform error) in the following sense.  Let $\chi$ be an arbitrary error distribution.  Define $\chi'$ as the distribution where $m$ dimensions are sampled according to $\chi$ and the remaining dimensions have some fixed error.  Then, security of $\distLWE_{n, m, q, \chi}$ implies security of $\distLWE_{n+\alpha+ \beta, m+\alpha, q, \chi'}$.

@@

\begin{figure}[p]
\begin{framed}
\begin{enumerate}
\item Input $\vA', \vb'$, where $\vA' \overset{\$} \leftarrow \Fq^{m\times n}$ and $\vb'$ is either uniform over $\Fq^m$ or $\vb' = \vA'\vx' +\ve'$.
\item Randomly generate $\vect{R} \overset{\$}\leftarrow \Fq^{\alpha \times n}$. Initialize $\vQ \in \Fq^{n\times (\alpha+\beta)}$ to be the zero matrix.
\item Let $\vb^* = (\vb', b^*_{m+1}, \ldots,b^*_{m+\alpha})$, where $(b^*_{m+1}, \ldots, b^*_{m+\alpha} )\overset{\$} \leftarrow \Fq^\alpha$.\label{step:b generation}
\item Randomly generate $\vect{S} \overset{\$}\leftarrow \Fq^{\alpha \times (\alpha+\beta)}$.
		\subitem If $\rank(\vect{S})<\alpha$, stop and output a random bit.
\item Find a set of $\alpha$ linearly independent columns in $\vS$.  Let $T$ be the set of these indices.
\item For all $x_i$ where $n\leq i$ and $i\not \in T$:\label{step:fill in matrix}
\subitem Generate $x_i\overset{\$}\leftarrow \Fq$.  
\subitem For $j=1,..., m$:
\subsubitem Generate $\vQ_{j, i}\overset{\$}\leftarrow \Fq$
\subsubitem Set $b_j^* = b_j^* + \vQ_{j, i} x_i$.
\item Initialize $\vA^*  = \left(\begin{array}{c | c}\vA' & \vQ\\\hline \vR & \vS\end{array}\right)$.
\item \label{step:randomization}
For {$i=1,..., m$}:
\subitem Randomly generate $\gamma_i \leftarrow \Fq^{1 \times \alpha}$.
\subitem Set $\va_{i} \leftarrow \va^*_{i}+\gamma_i (\vR||\vS)$
\subitem Set $b_i \leftarrow b^*_i + \gamma_i (b^*_{m+1},..., b^*_{m+k})$
\item Run $D$ on input $\vA, \vb$.
\item Output $D$'s output.
\end{enumerate}
\end{framed}
\caption{A PPT $D'$ that distinguishes LWE using distinguisher for LWE w/ block fixing source}
\label{fig:perfectLWEreduction}
\end{figure}
The distinguisher $D'$ has an advantage when $S$ is of rank $\alpha$.  This occurs with overwhelming probability~(techniques from Cooper~\cite{cooper2000rank}):
\begin{claim}
\label{cl:full rank matrix}
Let $\vS \overset{\$}\leftarrow \Fq^{\alpha \times (\alpha+\beta)}$ be randomly generated.  Then $\Pr[\rank(\vS)=\alpha]\geq  1- \ngl(n)$.
\end{claim}
\begin{proof}
The probability that all $\alpha$ rows are linearly independent is the successive probability that each row is linearly independent of the previous rows, for the $i$th row this probability is $q^{i-1}/q^{\alpha+\beta}$.  By union bound and $\beta = \omega(1)$, the probability that $\vS$ is not full rank is:
\begin{align*}
\Pr[\rank(\vS) < \alpha] &= \sum_{i=1}^{\alpha} \frac{q^{i-1}}{q^{\alpha+\beta}}
\leq \alpha \frac{q^{\alpha}}{q^{\alpha+\beta} }=\ngl(n)\\
\end{align*}
Thus, $\Pr[\rank(\vS) = \alpha] \geq 1-\ngl(n)$.
\end{proof}
\begin{corollary}
$
|\Pr[D(\vA, \vA\vx+\ve) = 1  | \rank(\vS) = \alpha] - \Pr[D(\vA, \vu) =1 | \rank(\vS) = \alpha]|> \epsilon - \ngl(n).
$
\end{corollary}

@@

\begin{claim}
\label{cl:randomMatrixDist}
The matrix $\vA$ generated after Step \ref{step:randomization} is a random matrix with $\rank(\vS) = \alpha$.
\end{claim}
\begin{proof}
It should be clear that $\vR, \vS$ are randomly generated~(conditioned on $\rank(\vS) =\alpha$).  The quadrant represented by $\vA'$ remains random after step~\ref{step:randomization} as we are adding uncorrelated random values.  
The matrix $\vQ_{T^c}$~(recall this is the restriction of $\vQ$ to the columns not in $T$) is random as it is initialized with random values and only uncorrelated random values are added.

Thus, we restrict attention to $\vQ_T$~(the columns of $W$ included in $T$).  Consider a particular row $i$.  Then $\vQ_{i, T} = \gamma_i \vS_{T}$.  Since $\vS_T$ is full rank and $\gamma_i$ is uniformly generated, $\vQ_{i, T}$ is uniformly generated.

@@

Let $\vb^*$ be the vector generated after step~\ref{step:fill in matrix}.  The only correlation between $\vb^*$ and $\vA^*$ is in the matrix $\vQ$, i.e. for $1\leq i\leq n$, $b_i^* = b_i + \sum_{n<j \wedge j\not\in T} \vQ_{i, j} x_j$, thus conditioning on $\vQ$, $b_i^*$ is uniform assuming $b_i$ is uniform.  Thus $\vb^* | \vA^*$ is uniformly distributed. We show that $\vb |( \vA, \vA^*)$ is uniformly distributed, therefore, showing that $\vb | \vA$ is uniformly distributed after step~\ref{step:randomization}.  Let $\vgamma$ represent the matrix formed by $\gamma_{i}$.
